### 2018-10-21 appendix_s2.Rmd

---
title: 'Appendix S2: Forward algorithm description'
subtitle: 'From: **Disease and climate effects on individuals drive post-reintroduction population dynamics of an endangered amphibian** in the journal Ecosphere'
author: 'Maxwell B. Joseph & Roland A. Knapp'
output: 
  pdf_document
header-includes:
  - \newcommand{\matr}[1]{\mathbf{#1}}
---

Parameter estimation for this model is made somewhat difficult by the presence of discrete parameters (hidden states).
We address this issue by using the forward algorithm, which does not require sampling from the discrete state space, to compute the joint probability of hidden states and observations [@zucchini2016hidden].
To describe this algorithm, we first consider the case of one individual. 
We would like to compute $[u_{i, 1:n_t}, o_{i, 1:n_j} | ...]$ (suppressing dependence on detection and transition parameters for compactness) for the individual with state and capture history shown in Figure 1.  
We can factor this joint probability as follows:

$$[u_{i, 1:n_t}, o_{i, 1:n_j}] = [u_{i, 1}] [u_{i, 2} | u_{i, 1}] [u_{i, 3} | u_{i, 2}] [o_{i, j = 1} | u_{i, 3}] [o_{i, j = 2} | u_{i, 3}] [u_{i, 4} | u_{i, 3}] ... [u_{i, n_t} | u_{i, n_t - 1}] [o_{i, n_j} | u_{i, n_t}].$$

\noindent If all of the unknown states were known, this would be as simple as extracting the relevant probabilities from $\matr{\Psi}_{i, t}$ and $\matr{\Omega}_{i, j}$.
Assuming that all individuals in the first primary period are in the ``not recruited'' class ($u_{i, 1} = 1 \, \forall i$) implies that $[u_{i, 1}] = (1 \, 0 \, 0 \, 0)$, where each element in the row vector represents the probability of being in hidden state 1, 2, 3, and 4, respectively.
If we define: $\matr{P} (o_{i, j}) = \text{diag}(\matr{\Omega}_{., i, j, o_{i, j}})$ to be the square matrix acquired by placing the elements of column $o_{i, j}$ from $\matr{\Omega}_{i, j}$ along the diagonal (with zeros elsewhere), the forward algorithm provides the joint distribution of hidden states and observations as follows:

$$[u_{i, 1:n_t}, o_{i, 1:n_j}] = [u_{i, t=1}] \matr{\Psi}_{i, t=1} \matr{\Psi}_{i, t=2} \matr{P} (o_{i, j = 1}) \matr{P} (o_{i, j = 2}) \matr{\Psi}_{i, t=2} ... \matr{\Psi}_{i, t=n_t - 1}\matr{P} (o_{i, j = n_j}) \matr{1}'$$

\noindent where $\matr{1}'$ is a column vector of ones. 
More generally, we can compute this probability by defining $\matr{B}_{i, t} = \matr{\Psi}_{i, t - 1} \prod_{j \sim t} \matr{P} (o_{i, j})$, where $j \sim t$ indicates surveys that took place in primary period $t$ (if no surveys took place, then $\matr{B}_{i, t} = \matr{\Psi}_{i, t - 1}$).
Then bringing back dependence on all other detection and transition parameters into our notation, we can compute the joint probability of hidden states and the observation history compactly as:

$$[u_{i, 1:n_t}, o_{i, 1:n_j} | \theta^{(\lambda)}, \theta^{(\gamma)}, \theta^{(\phi)}, \theta^{(\eta)}, \theta^{(p)}, z_{i, 1:n_t}] = [u_{i, t=1}] \prod_{t  = 2}^{n_t}\matr{B}_{i, t} \matr{1}'.$$